Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increase Repair Time?

### 2015-03-14 phaseDelay.tex
upgraded pickture of Royce case study
small edits

\conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increase  Repair Time?}
\numberofauthors{3}
\author{
\alignauthor
Tim Menzies, \\Carter Pape\\
       \affaddr{CS, NcState, USA}\\
       tim.menzies@gmail.com,\\carterpape@gmail.com
\alignauthor
William Nichols,\\ Forrest Schull\\
\affaddr{SEI, CMU, USA}
wrn,fjshull@sei.cmu.edu
\alignauthor
Lucas \\Layman\\
       \affaddr{Fraunhofer Center,  USA}\\ 
       llayman@fc-md.umd.edu
}

\maketitle
\begin{abstract}
Does
repair time increases dramatically
the longer a defect persists in a system?
This  is very widely belief that is used to justify 
radical changed to    traditional software development
 processes.

This paper shows that this belief is maintained
quite strongly in the industrial software engineering
community (but, in the academic community, somewhat less so).
Yet based on a sample of 
212 software projects from around the world from 
\bill{2005 to 2014}, we argue that this belief is mostly 
incorrect. Specifically: the median fix time for issues
is just a few minutes and this does not tend to increase
the longer the issue remains in the system.

@@

\section{Introduction}
In 2013-2014, the time of 
eleven  million programmers~\cite{pettey14} and
half a trillion dollars~\cite{avram14} was spent on information technology.
Such a large and growing effort should be managed and optimized via  well-researched conclusions.  
It is standard practice
in other fields such as medicine,
to continually revisit old conclusions~\cite{prasad13}. Software engineering should do the same lest
ill-founded  exert an undue influence on how we build software.

Accordingly, this paper revisits
the {\em phase delay effect}
i.e.,  the time required to resolve an issue in a software project increases dramatically 
the longer it remains in the system. 
Phase delay has long been
a    motivator for major changes to
SE practices. Basili \& Boehm comment that since the 1980s,  phase delay:
\begin{quote}
...has been a major driver in focusing
industrial software practice on thorough
requirements analysis and design,
on early verification and validation, and
on up-front prototyping and simulation
to avoid costly downstream fixes~\cite{boehm01}.
\end{quote}
Basili and Boehm are not alone in this assessment of the significance of the phase delay effect. Our surveys of industrial practitioners (shown in \S\ref{survey}) reveal the phase-delay effect to be
the single most strongly held belief amongst commercial SE engineers.
Previously~\cite{me08a}, we have constructed return-on-investement models
for a common practice amongst U.S. government contractors: 
independent verification \& validation. IVV has been criticized as
ineffective since IVV engineers, who many  unfamiliar with the details of the project, are required to pass judgement on the code.
Hence,  the IVV team may find the same or fewer bugs
as the developers,
On the other hand, IVV has been defended
as follows:  when the IVV team finds bugs earlier that the developers then, due to the phase delay effect, this saves the overall cost of the project. Note that the strength of this argument diminishes
if fixing later-phase issues is not significantly more expensive than fixing them earlier.

@@

\bi
    \item $H_1$: The cost-to-fix of defects monotonically increases across phases.
    \item $H_2$: The cost-to-fix of defects monotonically increases as the phase delay increases.
\ei
The rest of the paper investigates these two hypotheses. First, we conduct a survey of practitioners and researchers to gauge their belief in the phase delay effect. Second, we  analyze 171 software  projects developed in the period \bill{2004 to 2014}.  To the best of our knowledge, this is the largest study on phase delay yet seen in the literature.

@@

\section{Motivation}
\subsection{Why Reassess Existing Triusms?}

@@

Belief in general
principles of software
engineering are common to both research and practice. Processional societies assume such generalities exist when they offer
 lists of supposedly general ``best practices'' such as
the IEEE 1012 standard for software verification~\cite{1012}.  Endres \& Rombach's comment that 
\begin{quote}
Based on repeated and consistent observations, key lessons of these fields can now be formulated into rules or even laws, providing initial building blocks towards a theoretical foundation that is essential for further research, for teaching and for the practice of software development~\cite{endres03,rombach11}
\end{quote}
 Endres \& Rombach offer dozens of lessons of software engineering~\cite{rombach11}.
 Many other 
widely-cited researchers  do the same; e.g.
Glass~\cite{glass02}; Jones~\cite{jones07}; Boehm~\cite{boehm00b}.
Budgen \& Kitchenham seek to reorganize SE research using
general
conclusions drawn from a larger number of studies~\cite{kitch04,budgen09}.

@@

\item
Ray et al.~\cite{ray2014lang} tested if   strongly typed languages
predict for better code quality. In  728 projects,
they found  only a modest benefit in strong typing
(and caution that that effect may actually be due to other  conflating factors).
\item
Fenton \& Neil~\cite{fenton00,fenton00b}   critique the truism that
``pre-release fault rates for software
are a predictor for post-release failures'' (as claimed by~\cite{dunsmore88},
amongst others). For the systems described in~\cite{fenton97}, they
show that software modules that were highly fault-prone
prior to release revealed very few faults after release.
\item
Meyer claims that   object-oriented (OO) encapsulation will
reduce error rates in software~\cite{Meyer1988}.  Yet empirical results suggest
that debugging an OO program is many times harder and
longer than debugging a standard procedural program~\cite{hatton98}.
\item A truism of visual programming is that ``visual
representations are inherently superior to mere textual representations''. A review by Menzies suggests that the available
evidence for this claim is hardly conclusive~\cite{me00v}. 
\item Numerous recent {\em local learning} results compare single models
learned from all available data to multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,betta12,posnett11}.
A repeated result in those studies in that the local models generated the better effort
and defect predictions (better median results,
lower variance in the predictions).
 \item 
Other work studies  found dozens of single
models that performed worse that {\em committees of models}, all with different opinions,
that contribute to an {\em ensemble} prediction of software development effort~\cite{kocaguneli2012value,azhar13}.
\ei
%To be fair to the software engineering community,
Software engineering is not the only
field where practitioners hang on to persistent beliefs, even if the evidence
for those beliefs is not strong.
The medical profession applies  many practices based on studies
that have been disproved. For example,
a  recent article
in the Mayo Clinic Proceedings~\cite{prasad13} found  
146 medical practices based on studies 
in year $i$, but which were  reversed by subsequent trials within years $i+10$.
Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}.
Aschwanden warns:
\begin{quote}
Cognitive biases - such as motivated reasoning (all of us want to believe that the things we do make a difference), base rate neglect (failing to pay attention to what happens in the absence of the intervention), and confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) - also influence how we process information~\cite{aschwanden15}.
\end{quote}

@@

In summary, just like medicine, our field suffers when
 software engineers do  revise old beliefs.  Therefore, it is important
 to regularly  reexamine    old beliefs.

@@

Beck speculated that there might be another kind of project, shown in \fig{curve2},
that does no exhibit phase delay. In this next quote, Beck clearly states
the importance of avoiding phase delay (in fact, for Beck, it is {\em the} major
motivation for his work):
\begin{quote}
This (\fig{curve2}) is one of the premises of (agile). It is the technical premise of (agile). If the cost of change rose slowly over time, you would act completely differently from how you do under the assumption that costs rise exponentially. You would make big decisions as late in the process as possible, to defer the cost of making the decisions and to have the greatest possible chance that they would be right. You would only implement what you had to, in hopes that the needs you anticipate for tomorrow wouldn't come true. You would introduce elements to the design only as they simplified existing code or made writing the next bit of code simpler.~\cite{beck00}
\end{quote}
Beck makes one other comment that is relevant to this paper: 
\begin{quote}
The software development community has spent enormous resources in recent decades trying to reduce the cost of change-better languages, better database technology, better programming practices, better environments and tools, new notations.

@@

XXX clumsy link here
That said, where our results
diverge from Beck is that we have observed this lack of phase delay in  projects that are not agile. Those
projects are described below.

@@

The distribution of defects found and fixed per phase is shown in Figure~\ref{fig:fix-phase-dist}. A high percentage of defects were found and fixed in the early phases, i.e., requirements, high level design, and design reviews and inspections. \todo{How does this compare evidence from other sources?}

@@

The relative cost to fix errors per lifecycle phase in the TSP project data is overlayed with Boehm's exponential cost curve~\cite{Boehm81} as well as the Royce case study~\cite{Royce98} in \fig{cost-to-fix-tsp}.

\begin{figure}[!ht]
 \includegraphics[width=3.3in]{boehm-overlay.pdf}
 \caption{Cost-to-fix curve from \fig{cost-to-fix} overlayed with TSP data and case study from~\cite{Royce98}.}
 \label{fig:cost-to-fix-tsp}
 \end{figure}

@@

@@

\fig{phase_delay_medians} and \fig{correlation} show that most defects have a relatively small phase delay in our project sample (phase delay is often <= 2). As intended, the Review and Inspection Phases for Requirements, Design, and Code mostly find defects originating in the preceding phase. In contrast, the Testing phases mostly find defects originating in the Code phase.

In \fig{raw}, we present the median fix times of defects based on when they were injected and ultimately removed. Defects injected Before Development follow the most severe scale up as development phases pass, but not on the exponential scale traditionally seen (\fig{cost-to-fix}). Defects injected in Requirements, Design, and Coding phases exhibit virtually no scale-up regardless of the phase in which they are fixed.

@@

\begin{figure*}[!t] 
 \renewcommand{\baselinestretch}{0.7}
 \scriptsize
\begin{center}
\begin{tabular}{r|rrr|ll|rl}
  Sample&\multicolumn{3}{c|}{Percentiles}\\ 
size & 25th & 50th & 75th & Phase injected & Phase removed & \multicolumn{2}{l}{Scale up w.r.t. to first phase}\\\hline
\\
  29 &     1 &     3 &    12 & Before & CodeInspect  & 1.00  &  **  \\
  26 &     1 &     4 &     6 & Before & CodeReview  & 1.33  &  ***  \\
  73 &     5 &    13 &    59 & Before & UnitTest  & 4.33  &  *********  \\
  36 &     7 &    24 &    54 & Before & IntTest  & 8.00  &  ****************  \\
  34 &     8 &    28 &    56 & Before & SysTest  & 9.33  &  *******************  \\\hline
\\
128 &     2 &     6 &    15 & Reqts & ReqtsInspect  & 1.00  &  **  \\
132 &     2 &     5 &    12 & Reqts & ReqtsReview  & 0.83  &  **  \\
 28 &     3 &    13 &    29 & Reqts & DesignInspect  & 2.16  &  *****  \\
 22 &     6 &    14 &    31 & Reqts & DesignReview  & 2.33  &  *****  \\\hline
\\
272 &     6 &    14 &    30 & Design & DesignInspect  & 1.00  &  **  \\
 32 &     1 &     4 &    13 & Design & DesignInspection  & 0.28  &  *  \\
240 &     4 &    11 &    22 & Design & DesignReview  & 0.78  &  **  \\
110 &     3 &    11 &    23 & Design & CodeInspect  & 0.78  &  **  \\
 91 &     3 &     7 &    18 & Design & CodeReview  & 0.50  &  *  \\
323 &     8 &    21 &    45 & Design & UnitTest  & 1.50  &  ***  \\
 56 &     7 &    20 &    41 & Design & IntTest  & 1.42  &  ***  \\
 69 &     9 &    27 &    53 & Design & SysTest  & 1.92  &  ****  \\\hline
\\
310 &     6 &    15 &    31 & Code & CodeInspect  & 1.00  &  **  \\
285 &     5 &    14 &    29 & Code & CodeReview  & 0.93  &  **  \\
287 &     7 &    17 &    44 & Code & UnitTest  & 1.13  &  ***  \\
180 &     7 &    20 &    45 & Code & IntTest  & 1.33  &  ***  \\
137 &     6 &    15 &    35 & Code & SysTest  & 1.00  &  **  \\
 37 &     1 &    11 &    27 & Code & AcceptTest  & 0.73  &  **  \\

@@

 \section{Conclusion}
 \todo{
 \bi
    \item Summarize data tested - survey + TSP data
    \item reiterate hypothesis testing results
    \item Restate the motivation: widely-held beliefs need to be revisted. 
    \item We now have evidence, from 5 different sources, that phase delay does not seem to hold under certain conditions (although our study is by far the biggest in terms of data points).
    \item phase delay may still apply in some (many?) projects, but clearly it does not in all. This has huge implications for, e.g., cost modeling, effort prediction.
 \ei
 }

@@

\section*{Acknowledgements}
This work was partially funded by an National Science
Foundation grants NSF-CISE 1302169 and CISE 1302169. 
\\
We thank Tuma Solutions for providing the Team Process Data Warehouse software in support of this research.

>>>>

\conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increase  Repair Time?}
\numberofauthors{3}
\author{
\alignauthor
Tim Menzies, \\Carter Pape\\
       \affaddr{Dept. of Computer Science\\
       North Carolina State U, USA}\\
       tim.menzies@gmail.com,\\carterpape@gmail.com
\alignauthor
William Nichols,\\ Forrest Shull\\
\affaddr{Software Engineering Institute\\
        Carnegie Mellon U, USA}
wrn,fjshull@sei.cmu.edu
\alignauthor
Lucas \\Layman\\
       \affaddr{Fraunhofer CESE, \\
                College Park, USA}\\ 
       llayman@fc-md.umd.edu
}

\maketitle
\begin{abstract}
Does
repair time increase dramatically
the longer a defect persists in a system?
This widely held belief is used to justify 
major investments in  software development
 processes and methods.

This paper shows that this belief is maintained
quite strongly in the industrial software engineering
community (but, in the academic community, somewhat less so).
Yet based on a sample of 
171 software projects from around the world from 
and completed between 2006 and 2014, we argue that this belief is mostly 
incorrect. Specifically: the median fix time for issues
is just a few minutes and this does not tend to increase
the longer the issue remains in the system.

@@

\section{Introduction}
In 2013-2014, the time of 
eleven  million programmers~\cite{pettey14} and
half a trillion dollars~\cite{avram14} was spent on information technology.
Such a large and growing effort should be managed and optimized via  well-researched conclusions.  
It is standard practice
in other fields such as medicine,
to continually revisit old conclusions~\cite{prasad13}. Software engineering should do the same lest
ill-founded guidance exert an undue influence on how we build software.

Accordingly, this paper revisits
the {\em phase delay effect}
i.e.,  the rule of thumb stating that the time required to resolve an issue in a software project increases dramatically 
the longer it remains in the system. 
Phase delay has long been
a    motivator for major changes to
SE practices. Basili \& Boehm comment that since the 1980s,  phase delay:
\begin{quote}
...has been a major driver in focusing
industrial software practice on thorough
requirements analysis and design,
on early verification and validation, and
on up-front prototyping and simulation
to avoid costly downstream fixes~\cite{boehm01}.
\end{quote}
Basili and Boehm are not alone in this assessment of the significance of the phase delay effect. Our surveys of industrial practitioners (shown in \S\ref{survey}) reveal the phase-delay effect to be
the single most strongly held belief amongst commercial SE engineers.
Previously~\cite{me08a}, we have constructed return-on-investement models
for a common practice amongst U.S. government contractors: 
independent verification \& validation. IVV has been criticized as
ineffective since IVV engineers, many of whom are unfamiliar with the details of the project, are required to pass judgement on the code.
Hence,  the IVV team may find the same or fewer bugs
as the developers.
On the other hand, IVV has been defended
as follows:  when the IVV team finds bugs earlier than the developers then, due to the phase delay effect, this provides overall cost savings to the project. Note that the strength of this argument diminishes
if fixing later-phase issues is not significantly more expensive than fixing them earlier.

@@

\bi
    \item $H_1$: The cost-to-fix of defects monotonically increases across phases.
    \item $H_2$: The cost-to-fix of defects monotonically increases as the phase delay increases.
\ei
The rest of the paper investigates these two hypotheses. First, we conduct a survey of practitioners and researchers to gauge their belief in the phase delay effect. Second, we  analyze 171 software  projects developed in the period 2006 to 2014.  To the best of our knowledge, this is the largest study on phase delay yet seen in the literature.

@@

\section{Motivation}
\subsection{Why Reassess Existing Truisms?}

@@

Belief in general
principles of software
engineering are common to both research and practice. Professional societies assume such generalities exist when they offer
 lists of supposedly general ``best practices'' such as
the IEEE 1012 standard for software verification~\cite{1012}.  Endres \& Rombach's comment that 
\begin{quote}
Based on repeated and consistent observations, key lessons of these fields can now be formulated into rules or even laws, providing initial building blocks towards a theoretical foundation that is essential for further research, for teaching and for the practice of software development~\cite{endres03,rombach11}
\end{quote}
 Endres \& Rombach offer dozens of lessons of software engineering~\cite{rombach11}.
 Many other 
widely-cited researchers  do the same; e.g.
Glass~\cite{glass02}; Jones~\cite{jones07}; Boehm~\cite{boehm00b}.
Budgen \& Kitchenham seek to reorganize SE research using
general
conclusions drawn from a larger number of studies~\cite{kitch04,budgen09}.

@@

\item
Ray et al.~\cite{ray2014lang} tested if   strongly typed languages
predict for better code quality. In  728 projects,
they found  only a modest benefit in strong typing
(and caution that that effect may actually be due to other  conflating factors).
\item
Fenton \& Neil~\cite{fenton00,fenton00b}   critique the truism that
``pre-release fault rates for software
are a predictor for post-release failures'' (as claimed by~\cite{dunsmore88},
amongst others). For the systems described in~\cite{fenton97}, they
show that software modules that were highly fault-prone
prior to release revealed very few faults after release.
\item
Meyer claims that   object-oriented (OO) encapsulation will
reduce error rates in software~\cite{Meyer1988}.  Yet empirical results suggest
that debugging an OO program is many times harder and
longer than debugging a standard procedural program~\cite{hatton98}.
\item A truism of visual programming is that ``visual
representations are inherently superior to mere textual representations''. A review by Menzies suggests that the available
evidence for this claim is hardly conclusive~\cite{me00v}. 
\item Numerous recent {\em local learning} results compare single models
learned from all available data to multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,betta12,posnett11}.
A repeated result in those studies in that the local models generated the better effort
and defect predictions (better median results,
lower variance in the predictions).
% \item 
% Other work studies  found dozens of single
% models that performed worse that {\em committees of models}, all with % different opinions,
% that contribute to an {\em ensemble} prediction of software 
% development effort~\cite{kocaguneli2012value,azhar13}.
\ei
%To be fair to the software engineering community,
Software engineering is not the only
field where practitioners hang on to persistent beliefs, even if the evidence
for those beliefs is not strong.
The medical profession applies  many practices based on studies
that have been disproved. For example,
a  recent article
in the Mayo Clinic Proceedings~\cite{prasad13} found  
146 medical practices based on studies 
in year $i$, but which were  reversed by subsequent trials within years $i+10$.
Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}.
Aschwanden warns:
\begin{quote}
Cognitive biases - such as motivated reasoning (all of us want to believe that the things we do make a difference), base rate neglect (failing to pay attention to what happens in the absence of the intervention), and confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) - also influence how we process information~\cite{aschwanden15}.
\end{quote}

@@

In summary, just like medicine, our field suffers when
 software engineers do  not revise old beliefs.  Therefore, it is important
 to regularly  reexamine    old beliefs.

@@

Beck speculated that there might be another kind of project, shown in \fig{curve2},
that does not exhibit phase delay. In this next quote, Beck clearly states
the importance of avoiding phase delay (in fact, for Beck, it is {\em the} major
motivation for his work):
\begin{quote}
This (\fig{curve2}) is one of the premises of (agile). It is the technical premise of (agile). If the cost of change rose slowly over time, you would act completely differently from how you do under the assumption that costs rise exponentially. You would make big decisions as late in the process as possible, to defer the cost of making the decisions and to have the greatest possible chance that they would be right. You would only implement what you had to, in hopes that the needs you anticipate for tomorrow wouldn't come true. You would introduce elements to the design only as they simplified existing code or made writing the next bit of code simpler.~\cite{beck00}
\end{quote}
Beck makes one other comment that is relevant to this paper: 
\begin{quote}
The software development community has spent enormous resources in recent decades trying to reduce the cost of change-better languages, better database technology, better programming practices, better environments and tools, new notations.

@@

%XXX clumsy link here
%That said, where our results
%diverge from Beck is that we have observed this lack of phase delay in  projects that are not agile. Those
%projects are described below.

@@

Our first hypothesis was that defects found in later phases tend to require more time to fix than in earlier phases.

The distribution of defects found and fixed per phase in our data is shown in Figure~\ref{fig:fix-phase-dist}. A high percentage of defects were found and fixed in the early phases, i.e., requirements, high level design, and design reviews and inspections. \todo{How does this compare evidence from other sources?}

@@

For illustrative purposes, the relative cost to fix errors per lifecycle phase in the TSP project data is overlayed with Boehm's cost curve~\cite{Boehm81} as well as the CCPDS-R case study~\cite{Royce98} in \fig{cost-to-fix-tsp}. It is clear to see that the trend is relatively flat, closer to the pattern exhibited by the Royce case study (which also focused on early defect removal) than the exponential growth curve.

\begin{figure}[!ht]
 \includegraphics[width=3.3in]{boehm-overlay.pdf}
 \caption{Cost-to-fix curve from \fig{cost-to-fix} overlayed with case study from~\cite{Royce98} and TSP data. Scale is relative to the phase with the lowest cost to fix. }
 \label{fig:cost-to-fix-tsp}
 \end{figure}

@@

Our second hypothesis was that defects that exhibit higher phase delay (i.e. that persist in the software over more lifecycle phases) tend to require more effort to fix than those with less phase delay. To investigate this, we calculate phase delay as simply the number of project phases (the set used in our data set is shown in \fig{waterfall}) during which the defect existed in the system, i.e. the ordinal rank of removal phase minus the ordinal rank of injection phase. Sub-phases (such as review and inspection) are not included in this calculation.

\fig{correlation} and \fig{phase_delay_medians} show that most defects have a relatively small phase delay in our project sample (phase delay is often <= 2). As intended, the Review and Inspection Phases for Requirements, Design, and Code mostly find defects originating in the same phase. In contrast, the Testing phases mostly find defects originating in the Code phase.

In \fig{raw}, we present the median fix times of defects based on when they were injected and ultimately removed. Defects injected Before Development (labelled "B4Dev" in \fig{waterfall}) follow the most severe scale up as development phases pass, but not on the exponential scale traditionally seen (\fig{cost-to-fix}). Defects injected in Requirements, Design, and Coding phases exhibit virtually no scale-up regardless of the phase in which they are fixed.

@@

@@

\begin{figure*}[!t] 
 \renewcommand{\baselinestretch}{0.7}
 \scriptsize
\begin{center}
\begin{tabular}{r|rrr|ll|rl}
  Sample&\multicolumn{3}{c|}{Percentiles}\\ 
size & 25th & 50th & 75th & Phase injected & Phase removed & \multicolumn{2}{l}{Scale up w.r.t. to first phase}\\\hline
\\
  29 &     1 &     3 &    12 & Before Dev & CodeInspect  & 1.00  &  **  \\
  26 &     1 &     4 &     6 & Before Dev & CodeReview  & 1.33  &  ***  \\
  73 &     5 &    13 &    59 & Before Dev & UnitTest  & 4.33  &  *********  \\
  36 &     7 &    24 &    54 & Before Dev & IntTest  & 8.00  &  ****************  \\
  34 &     8 &    28 &    56 & Before Dev & SysTest  & 9.33  &  *******************  \\\hline
\\
128 &     2 &     6 &    15 & Reqts & ReqtsInspect  & 1.00  &  **  \\
132 &     2 &     5 &    12 & Reqts & ReqtsReview  & 0.83  &  **  \\
 28 &     3 &    13 &    29 & Reqts & DesignInspect  & 2.16  &  *****  \\
 22 &     6 &    14 &    31 & Reqts & DesignReview  & 2.33  &  *****  \\\hline
\\
272 &     6 &    14 &    30 & Design & DesignInspect  & 1.00  &  **  \\
 32 &     1 &     4 &    13 & Design & DesignInspection  & 0.28  &  *  \\
240 &     4 &    11 &    22 & Design & DesignReview  & 0.78  &  **  \\
110 &     3 &    11 &    23 & Design & CodeInspect  & 0.78  &  **  \\
 91 &     3 &     7 &    18 & Design & CodeReview  & 0.50  &  *  \\
323 &     8 &    21 &    45 & Design & UnitTest  & 1.50  &  ***  \\
 56 &     7 &    20 &    41 & Design & IntTest  & 1.42  &  ***  \\
 69 &     9 &    27 &    53 & Design & SysTest  & 1.92  &  ****  \\\hline
\\
310 &     6 &    15 &    31 & Code & CodeInspect  & 1.00  &  **  \\
285 &     5 &    14 &    29 & Code & CodeReview  & 0.93  &  **  \\
287 &     7 &    17 &    44 & Code & UnitTest  & 1.13  &  ***  \\
180 &     7 &    20 &    45 & Code & IntTest  & 1.33  &  ***  \\
137 &     6 &    15 &    35 & Code & SysTest  & 1.00  &  **  \\
 37 &     1 &    11 &    27 & Code & AcceptTest  & 0.73  &  **  \\

@@

 \section{Conclusion}
 \todo{
 \bi
    \item Summarize data tested - survey + TSP data
    \item reiterate hypothesis testing results
    \item Restate the motivation: widely-held beliefs need to be revisited. 
    \item We now have evidence, from 5 different sources, that phase delay does not seem to hold under certain conditions (although our study is by far the biggest in terms of data points).
    \item phase delay may still apply in some (many?) projects, but clearly it does not in all. This has huge implications for, e.g., cost modeling, effort prediction.
 \ei
 }

@@

\section*{Acknowledgements}
This work was partially funded by an National Science
Foundation grants NSF-CISE 1302169 and CISE 1302169. 
\\
We thank Tuma Solutions for providing the Team Process Data Warehouse software in support of this research.  \\
Team Software ProcessSM and TSPSM are service marks of Carnegie Mellon University.
